---
title: "p8105_hw2_sl4836"
author: "Hun"
date: "10/5/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 0
```{r, warning = FALSE, message=FALSE}
getwd()

dir.create(file.path(getwd(), "local_data_file"), recursive = TRUE)

list.files()
```

## Problem 1 (Trash Wheel Data)
```{r, message=FALSE}

clean_trash_data <- read_excel(
  "local_data_file/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
  sheet = "Mr. Trash Wheel", range = "A2:N408") %>%
  janitor::clean_names() %>%
  filter(!str_detect(month, 'Total')) %>% 
  mutate(sports_balls = round(sports_balls))

clean_trash_data 
```

## Problem 1 (Precipitation 2018, 2019 Data)
```{r}
clean_precipation_2018_data <- read_excel(
  "local_data_file/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
  sheet = "2018 Precipitation",
  range = "A2:B15") %>% janitor::clean_names() %>% mutate(year = "2018") %>% 
  relocate(year) %>%
  drop_na()

clean_precipation_2019_data <- read_excel(
  "local_data_file/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
  sheet = "2019 Precipitation",
  range = "A2:B15") %>% janitor::clean_names() %>% mutate(year = "2019") %>%
  relocate(year) %>%
  drop_na()

combined_precipitation <- bind_rows(clean_precipation_2018_data, 
                                    clean_precipation_2019_data) %>%
  mutate(month = month.name[month]) %>% 
  relocate(year)

combined_precipitation
```
# Data Summarization of thef Problem_1 Datasets
```{r}
trash_variables <- names(clean_trash_data)
trash_nrow <- nrow(clean_trash_data)
trash_ncol <- ncol(clean_trash_data)
median_sports_balls <- clean_trash_data %>% 
  filter(year == 2019) %>% 
  summarise(median(sports_balls)) 
mean_sports_balls <- clean_trash_data %>% 
  filter(year == 2019) %>%  summarise(mean(sports_balls))
  

prec_2018_variables <- names(clean_precipation_2018_data)
prec_2018_nrow <- nrow(clean_precipation_2018_data)
prec_2018_ncol <- ncol(clean_precipation_2018_data)
total_precip_2018 <- clean_precipation_2018_data %>% summarise(sum(total))


prec_2019_variables <- names(clean_precipation_2019_data)
prec_2019_nrow <- nrow(clean_precipation_2019_data)
prec_2019_ncol <- ncol(clean_precipation_2019_data)
total_precip_2019 <- clean_precipation_2019_data %>% summarise(sum(total))

prec_combined_variables <- names(combined_precipitation)
prec_combined_nrow <- nrow(combined_precipitation)
prec_combined_ncol <- ncol(combined_precipitation)
total_combined_precip <- combined_precipitation %>% summarise(sum(total))



```

The dimension of the trash_data is **`r trash_nrow` x**  **`r trash_ncol`.** There are **`r trash_nrow`** number of observations and **`r trash_ncol`**  variables: *`r trash_variables`.* The median number of sports balls in a dumpster in 2019 is **`r median_sports_balls`.** The mean number of sports balls in a dumpster in 2019 is **`r mean_sports_balls`** 

The dimension of the 2018 precipitation data is **`r prec_2018_nrow` x**  **`r prec_2018_ncol`.** There are **`r prec_2018_nrow`** number of observations and **`r prec_2018_ncol`**  variables: *`r prec_2018_variables`.* The total precipitation in 2018 is **`r total_precip_2018`**

The dimension of the 2019 precipitation data is **`r prec_2019_nrow` x**  **`r prec_2019_ncol`.** There are **`r prec_2019_nrow`** number of observations and **`r prec_2019_ncol`**  variables: *`r prec_2019_variables`.* The total precipitation in 2019 is **`r total_precip_2019`**

The dimension of the combined precipitation data is **`r prec_combined_nrow` x**  **`r prec_combined_ncol`.** There are **`r prec_combined_nrow`** number of observations and **`r prec_combined_ncol`**  variables: *`r prec_combined_variables`.* The total precipitation of the combined 2018 and 2019 is **`r total_combined_precip`**



## Problem 2 (1 - Pols Month Data)
```{r, message=FALSE}
pols_month_data <- read_csv("local_data_file/pols-month.csv")

clean_pols_month_data <- pols_month_data %>%
  janitor::clean_names() %>% 
  separate("mon", c("year", "month", "day"), sep = "-") %>% 
  mutate(across(.cols = c("year", "month", "day"), as.integer)) %>% 
  mutate(month = month.name[month])  %>% 
  mutate(president = case_when(prez_dem == 1 ~ "dem",
                               prez_gop == 1 ~ "gop",
                               prez_gop == 2 ~ "gop")) %>%
    select(-day, -prez_dem, -prez_gop)
  

clean_pols_month_data

```

## Problem 2 (2 - Snp Data)
```{r, message=FALSE}
snp_data <- read_csv("local_data_file/snp.csv")

clean_snp_data <- snp_data %>% 
  separate("date", c("month", "day", "year"), sep = "/") %>%
  mutate(across(.cols = c("year", "month", "day"), as.integer)) %>%
  relocate(year, month) %>%
  mutate(month = month.name[month]) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year<21, year+2000, year+1900)) %>% 
  select(-day)

clean_snp_data
```

## Problem 2 (3 - Unemployment Data)
```{r, message=FALSE}
unemployment_data <- read_csv("local_data_file/unemployment.csv")

clean_unemployment_data <- unemployment_data %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment_percent") %>% 
  janitor::clean_names() %>% 
  mutate(across(.cols = c("year"), as.integer))  %>%
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month])

clean_unemployment_data
```
#Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r, message=FALSE}
merge1 <- left_join(clean_pols_month_data, clean_snp_data, by = c("year", "month"))

merge2 <- left_join(merge1, clean_unemployment_data, by = c("year", "month"))

merge2
```
## short paragraph about the Problem_2 Datasets
```{r, message=FALSE}
name_variables <- names(pols_month_data)
nrow <- nrow(pols_month_data)
ncol <- ncol(pols_month_data)

name_variables1 <- names(clean_pols_month_data)
nrow1 <- nrow(clean_pols_month_data)
ncol1 <- ncol(clean_pols_month_data)
range_year1 <- clean_pols_month_data %>% pull(year) %>% range()

name_variables2 <- names(snp_data)
nrow2 <- nrow(snp_data)
ncol2 <- ncol(snp_data)

name_variables3 <- names(clean_snp_data)
nrow3 <- nrow(clean_snp_data)
ncol3 <- ncol(clean_snp_data)
range_year3 <- clean_snp_data %>% pull(year) %>% range()

name_variables4 <- names(unemployment_data)
nrow4 <- nrow(unemployment_data)
ncol4 <- ncol(unemployment_data)

name_variables5 <- names(clean_unemployment_data)
nrow5 <- nrow(clean_unemployment_data)
ncol5 <- ncol(clean_unemployment_data)
range_year5 <- clean_unemployment_data %>% pull(year) %>% range()

name_variables6 <- names(merge1)
nrow6 <- nrow(merge1)
ncol6 <- ncol(merge1)
range_year6 <- merge1 %>% pull(year) %>% range()

name_variables7 <- names(merge2)
nrow7 <- nrow(merge2)
ncol7 <- ncol(merge2)
range_year7 <- merge2 %>% pull(year) %>% range()

```
Originally, the dimension of the pols_month_data was **`r nrow` x**  **`r ncol`** and there are **`r ncol`** variables: *`r name_variables`.* After tidying it, the dimension is **`r nrow1` x**  **`r ncol1`.** and there are **`r ncol1`** variables: *`r name_variables1`.* The range of years is **[`r range_year1`].**

Originally, the dimension of the _data was **`r nrow2` x**  **`r ncol2`** and there are **`r ncol2`** variables: *`r name_variables2`.* After tidying it, the dimension is **`r nrow3` x**  **`r ncol3`.** and there are **`r ncol3`** variables: *`r name_variables3`.* The range of years is **[`r range_year3`].**

Origianlly the dimension of the pols_month_data was **`r nrow4` x**  **`r ncol4`** and there are **`r ncol4`** variables: *`r name_variables4`.* After tidying it, the dimension is **`r nrow5` x**  **`r ncol5`.** and there are **`r ncol5`** variables: *`r name_variables5`.* The range of years is **[`r range_year5`].**

After joining the datasets by merging snp into pols, the dimension of the dataset is **`r nrow6` x**  **`r ncol6`** and there are **`r ncol6`** variables: *`r name_variables6`.* The range of years is **[`r range_year6`].**

After joining the datasets by additionally merging unemployment to the merged dataset, the dimension of the dataset is **`r nrow7` x**  **`r ncol7`** and there are **`r ncol7`** variables: *`r name_variables7`.* The range of years is **[`r range_year7`].**


## Problem 3 (Cleaning Baby_Names Data for Olivia)
```{r}
library(dplyr)

baby_names_data <-read_csv("local_data_file/Popular_Baby_Names.csv")

clean_baby_names_data <- baby_names_data %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_title(gender)) %>% 
  mutate(ethnicity = str_to_title(ethnicity)) %>%
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>%
  distinct() %>% 
  mutate(ethnicity = 
           recode(ethnicity, "White Non Hisp" = 
           "White Non Hispanic", 
           "Black Non Hisp" = "Black Non Hispanic", 
           "Asian And Paci" = "Asian And Pacific Islander"))%>%
  filter(str_detect(childs_first_name, 'olivia')) %>%
  filter(str_detect(gender, 'Female')) %>%
  arrange(year_of_birth)  

clean_baby_names_data
```

#Creating the Table That Shows the Rank of the Name Olivia Popularity by Ethnicity Over Time.
```{r, message=FALSE}
clean_baby_names_data %>% select(rank, year_of_birth, ethnicity,
         childs_first_name) %>%
  pivot_wider(names_from = childs_first_name, values_from = rank,
              names_glue = "olivia_rank") %>% 
  pivot_wider(names_from = year_of_birth, values_from
              =olivia_rank) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  knitr::kable(caption = "**Table1: The Rank (1st to 102nd) of the Name Olivia Popularity by Ethnicity Over Time**", align = "c", format = "pipe")
```

#Cleaning Baby_names Data for Male
```{r}
clean_baby_male_data <- baby_names_data %>% 
  janitor::clean_names() %>%
  mutate(ethnicity = str_to_title(ethnicity), gender = str_to_title(gender), childs_first_name = str_to_title(childs_first_name)) %>%
  filter(str_detect(gender, 'Male') & rank ==1) %>%
  arrange(year_of_birth) %>%
  mutate(ethnicity = 
           recode(ethnicity, 
           "White Non Hisp" = "White Non Hispanic", 
           "Black Non Hisp" = "Black Non Hispanic", 
           "Asian And Paci" = "Asian And Pacific Islander")) %>% 
  select(-count, -gender, -rank) %>%
  distinct()
```

#Creating the Table That Shows The Most Popular Name Among Male Children Over Time.
```{r}
library(kableExtra)
clean_baby_male_data  %>% 
  pivot_wider(names_from = year_of_birth, values_from =
                childs_first_name ) %>% 
  knitr::kable(caption = "**Table2: The Most Popular Male Names by Ethnicity Over Time**", align = "c", format = "pipe")  
```

#Creating a Scaterplot Showing the Number of Children with a Name (y axis) against the Rank in Popularity of That Name (x axis).
```{r}
  baby_names_data %>% 
  janitor::clean_names() %>% 
  distinct() %>%
  mutate(ethnicity = str_to_title(ethnicity), gender = str_to_title(gender)) %>% 
  mutate(ethnicity = 
           recode(ethnicity, 
           "White Non Hisp" = "White Non Hispanic", 
           "Black Non Hisp" = "Black Non Hispanic", 
           "Asian And Paci" = "Asian And Pacific Islander")) %>%
  filter(ethnicity == "White Non Hispanic", gender == "Male",
         year_of_birth == 2016) %>%
  group_by(childs_first_name) %>% 
  ggplot(aes(rank, count)) + geom_point() + 
  labs(title = "The Number of White Non-Hispanic Male Children's First Names Born in 2016 by Rank", x = "Rank in Popularity", 
       y="Count") + theme(plot.title = element_text(hjust = 0.4, size = 12))
 
```





